---
title: "DATA 607 - Project 3 - Analysis"
author: "Keith Colella"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
```

## Read in data

```{r}
# jobs <- read_csv('output/job_listings_skills.csv')
jobs_long <- read_csv('output/job_listings_skills_long.csv')
```

## Visualizations

```{r}
jobs_long %>%
  group_by(job_id, highest_ed) %>%
  summarize(jobs = n(), .groups = 'keep') %>%
  ggplot(aes(highest_ed)) +
  geom_bar()

jobs_long %>%
  group_by(job_id, salary_max) %>%
  summarize(jobs = n(), .groups = 'keep') %>%
  ggplot(aes(salary_max)) +
  geom_histogram(binwidth = 20000) +
  scale_x_continuous(label = scales::comma,
                     limits = c(0,500000))

jobs_long %>%
  group_by(job_id, years_exp) %>%
  summarize(jobs = n(), .groups = 'keep') %>%
  ggplot(aes(years_exp)) +
  geom_bar()

jobs_long %>%
  group_by(job_id, years_exp, highest_ed) %>%
  summarize(jobs = n(), .groups = 'keep') %>%
  filter(!is.na(highest_ed)) %>%
  ggplot(aes(years_exp, ..density..)) +
  geom_histogram(binwidth = 3) +
  facet_grid(~highest_ed,
             labeller = label_wrap_gen(width=2, multi_line=TRUE))

jobs_long %>%
  group_by(job_id, salary_max, highest_ed) %>%
  summarize(jobs = n(), .groups = 'keep') %>%
  filter(!is.na(highest_ed)) %>%
  ggplot(aes(salary_max, ..density..)) +
  geom_histogram(binwidth = 20000) +
  scale_x_continuous(label = scales::comma,
                     limits = c(0,500000)) +
  theme(axis.text.x = element_text(angle=90)) +
  facet_grid(~highest_ed,
             labeller = label_wrap_gen(width=2, multi_line=TRUE))

## We need to do some cleanup on location
jobs_long %>%
  ggplot(aes(fct_rev(fct_infreq(location)))) +
  geom_bar() +
  coord_flip()
```

## Most Frequently Cited Skills

```{r}
jobs_long %>%
  group_by(skills) %>%
  summarize(citations = n(), .groups = 'keep') %>%
  arrange(desc(citations)) %>%
  head(20)

for (ed_level in unique(jobs_long$highest_ed)) {
  jobs_long %>%
    filter(highest_ed == ed_level) %>%
    group_by(skills) %>%
    summarize(citations = n(),
              ed_level = ed_level,
              .groups = 'keep') %>%
    arrange(desc(citations)) %>%
    head(10) %>%
    print()
}

for (loc in c('New York', 'Remote', 'France')) {
  jobs_long %>%
    filter(location == loc) %>%
    group_by(skills) %>%
    summarize(citations = n(),
              location = loc,
              .groups = 'keep') %>%
    arrange(desc(citations)) %>%
    head(10) %>%
    print()
}
```


Group the data by the skills column and count the number of counts for each skill:

```{r}
skill_counts <- jobs_long %>%
  group_by(skills) %>%
  summarize(counts = n(), .groups = 'keep')

```


Filter out any skills that you don't want to include in the plot

```{r}
overweighted <- c("data", "analyst", "analysis", "analytics","big data","business analytics", "data")
skill_counts <- skill_counts[!grepl(paste(overweighted, collapse="|"), skill_counts$skills), ]

```


Sort the data by the number of citations in descending order and select the top 50 most cited skills

```{r}
skill_counts <- skill_counts %>%
  arrange(desc(counts)) %>%
  head(50)

```



Create the plot using ggplot2 and ggwordcloud packages

```{r}
library(ggplot2)
library(ggwordcloud)

ggplot(skill_counts, aes(label = skills, size = counts, color = factor(sample.int(8, nrow(skill_counts), replace = TRUE)))) +
  geom_text_wordcloud() +
  scale_size_area(max_size = 10) +
  theme_minimal()

```

Most frequently counted skills for job listings with at least 5 years of experience

```{r}
exp5_skill_counts <- jobs_long %>%
  filter(years_exp >= 5) %>%
  group_by(skills) %>%
  summarize(counts = n(), .groups = 'keep')

ggplot(exp5_skill_counts, aes(label = skills, size = counts, color = factor(sample.int(8, nrow(exp5_skill_counts), replace = TRUE)))) +
  geom_text_wordcloud() +
  scale_size_area(max_size = 15) +
  theme_minimal()

```


```{r}
ggplot(jobs_long, aes(x = years_exp, y = salary_max, color = highest_ed)) +
  geom_point(alpha = 0.6) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = 'Years of experience', y = 'Maximum salary', color = 'Education level')

```



```{r}
library(ggplot2)
library(ggwordcloud)

# Plot a histogram of salaries by highest education level
ggplot(jobs_long, aes(x = salary_max, fill = highest_ed)) +
  geom_histogram(binwidth = 20000, alpha = 0.7) +
  scale_x_continuous(label = scales::comma, limits = c(0, 500000)) +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  labs(title = "Distribution of maximum salaries by highest education level",
       x = "Salary (USD)", y = "Count")

# Plot a scatterplot of salary versus years of experience, colored by location
ggplot(jobs_long, aes(x = years_exp, y = salary_max, color = location)) +
  geom_point(alpha = 0.7) +
  scale_x_continuous(limits = c(0, 20)) +
  scale_y_continuous(label = scales::comma, limits = c(0, 500000)) +
  labs(title = "Salary versus years of experience, by location",
       x = "Years of experience", y = "Salary (USD)")

# Plot a word cloud of the most frequently cited skills
skills_freq <- jobs_long %>%
  count(skills) %>%
  arrange(desc(n)) %>%
  head(100)
ggplot(skills_freq, aes(label = skills, size = n)) +
  geom_text_wordcloud() +
  theme_minimal() +
  labs(title = "Most frequently cited skills")

# Plot a bar chart of the number of job listings per website
ggplot(jobs_long, aes(x = website)) +
  geom_bar(fill = "steelblue") +
  coord_flip() +
  labs(title = "Number of job listings per website",
       x = "Website", y = "Count")

# Plot a stacked bar chart of the number of job listings by location and highest education level
ggplot(jobs_long, aes(x = highest_ed, fill = location)) +
  geom_bar() +
  labs(title = "Number of job listings by location and highest education level",
       x = "Highest education level", y = "Count") +
  theme(legend.position = "bottom")

```



